---
title: 'Alaska Paleofire Synthesis: Main'
author: "Tyler Hoecker"
date: "6/13/2018"
output: html_document
---
```{r}
library(dplyr)
library(purrr)
library(readr)
library(tidyr)
library(ggplot2)
library(broom)
library(zoo)
library(truncnorm)
```

## Composite indices of biomass burning 

Define lakes for biomass burning analysis.
```{r}
# Noatak River Basin (data from Higuera et al. 2011)
noatak <- c('LI','RA','PO','UC')  
# Brooks Range (data from Higuera et al. 2007,2009)
brooks <- c('CO','LC','RP','WK','XI') 
# Yukon Flats (data from Kelly et al. 2014)
yukon  <- c('CP','EP','GA','JA','LD','LT','LU','NR','PI','RE','RO','SL','WC','WI') 
# Copper River Basin (data from Barett et al. 2013)
copper <- c('CR','HD','MP1','SC') 

alaska <- c(noatak,brooks,yukon,copper)
```

Master function 
```{r}
source(file.path('code','ZIL_Run.r'))
```

Run for Alaska and all regions using specificed bandwidths (half the width of the moving window, within which the ZIL distribution will be estimated)
```{r}
bandWidth   <- 50 
output100 <- map(list(alaska, noatak, brooks, yukon, copper),
                   zil_fn)

compCHAR_100 <- output100 %>% 
  `names<-` (c('Alaska','Noatak','Kobuk','Yukon Flats','Copper')) %>%
  bind_rows(.id = 'region')

bandWidth   <- 250 	
output500 <- map(list(alaska, noatak, brooks, yukon, copper),
                   zil_fn) 

compCHAR_500 <- output500 %>% 
  `names<-` (c('Alaska','Noatak','Kobuk','Yukon Flats','Copper')) %>%
  bind_rows(.id = 'region')


compositeCHAR <- full_join(compCHAR_500, compCHAR_100, 
                           by = c('region', 'yr.bp'),
                           suffix = c('.500','.100'))
```

Quick plot of 100-year (thin line) and 500-year (thick line w/ CI) Alaska-wide and ecoregional composite indices of biomass burning. 
```{r}
ggplot() +
  geom_line(data = compCHAR_100, aes(x = yr.bp, y = composite.mean)) +
  geom_ribbon(data = compCHAR_500, aes(x = yr.bp, ymin = CI.lower, ymax = CI.upper), alpha = 0.3) +
  geom_line(data = compCHAR_500, aes(x = yr.bp, y = composite.mean), size = 1.25) +
  facet_wrap(~region) +
  scale_x_reverse() +
  coord_cartesian(ylim = c(0,3.5)) +
  theme_bw() +
  labs(x = 'Time (cal years BP)', y = 'Biomass burning (standardized CHAR)')
```

Re-create Figure 2, illustrating zero-inflated log-normal distribution of CHAR.
```{r}
# PLOT HERE
```

## Composite indices of fire frequency / fire synchrony

#### Percent sites burned per century

Redefine some regions (remove lakes with signal-to-noise index < 3)
```{r}
## Update ecoregion parameters
#  MP1 in Copper River  and XI in Kobuk not suitable for peak analysis
copper <- c("CR","SC","HD")
brooks <- c("CO","LC","RP","WK")

# Update 'regions' object
regions  <- list('noatak'= noatak,'brooks'= brooks,'yukon'= yukon, 'copper'= copper) %>% 
  stack() %>% 
  rename(lake = values, region = ind)
alaska <- c(noatak,brooks,yukon,copper)
```

Load synchrony (percent of sites burned per century) analysis, as a function, stored in Synchrony_Run.r script.
```{r}
source(file.path('code','synchrony_funs.r'))
```

Load the estimated timing of fire events from peak CHAR data.
```{r}
fireYears_df <- fireyears_fn(alaska)
```

Run the regional version of the synchrony analysis. This result is not presented in the main text of the manuscript.
```{r}
regional <- 1 #1 for yes, group by region, 0 for no, do all AK together
pctBurned_reg <- synch_fn(fireYears_df)
```

View those results:
```{r}
ggplot(pctBurned_reg, aes(x = year, y = boot.median)) +
  geom_step(size = 0.75, color = 'grey50', size = 0.5) +
  geom_smooth(aes(x = year, y = boot.upper), color = 'grey50', se = F, method = 'loess', span = 0.01, size = 0.5) +
  geom_smooth(aes(x = year, y = boot.lower), color = 'grey50', se = F, method = 'loess', span = 0.01, size = 0.5) +
  geom_smooth(size = 1, color = 'black', se = F, method = 'loess', span = 0.01) +
  theme_bw() +
  scale_x_reverse(limits = c(8000,0)) +
  labs(x = 'Time (cal years BP)', y = 'Synchrony: Percent sites burned per century') +
  facet_wrap(~region)
```

Run the Alaska-wide version of the synchrony analysis, as is presented in the manuscript. This portion of the code takes ~6 hours to run, so is commented out here, and loaded from an .RData file.
```{r}
regional <- 0 #1 for yes, group by region, 0 for no, do all AK together
pctBurned_ak <- synch_fn(fireYears_df)
```

View those results:
```{r}
ggplot(pctBurned_ak, aes(x = year, y = boot.median)) +
  geom_step(size = 0.75, color = 'grey50') +
  geom_smooth(aes(x = year, y = boot.upper), color = 'grey50', se = F, method = 'loess', span = 0.01) +
  geom_smooth(aes(x = year, y = boot.lower), color = 'grey50', se = F, method = 'loess', span = 0.01) +
  geom_smooth(size = 1, color = 'black', se = F, method = 'loess', span = 0.01) +
  theme_bw() +
  scale_x_reverse(limits = c(8000,0)) +
  labs(x = 'Time (cal years BP)', y = 'Synchrony: Percent sites burned per century')
```

#### Fire return intervals 

A 500-yr moving window mean estimate of FRI from peak-CHAR-estimated fire years was calcualted by PEH in MatLab. 
```{r}
FRIsmooth = read_csv(file.path('data','peakData','FRI_smooth.csv'))
FRIsmooth <- FRIsmooth %>% 
  filter(year_BP <= 8000)
```

View those results:
```{r}
# PLOT HERE
```

### Combine CHAR and FRI and calculate ratio of CHAR to FRI (i.e., fire severity)
```{r}
ratio <- compositeCHAR %>% 
  filter(region == 'Alaska') %>% 
  right_join(FRIsmooth, by = c('yr.bp' = 'year_BP')) %>%
  mutate(ratio = composite.mean.500/mFRI_1000,
         ratio.CIlower = CI.lower.500/CI_l,
         ratio.CIupper = CI.upper.500/CI_u) %>%
  select(yr.bp, ratio, ratio.CIlower, ratio.CIupper)
```

View those results:
```{r}
# PLOT HERE
```

## Climate data

#### Load climate data from Kaufman et al. (2017) and Wiles et al. (2014), and do some light munging.
```{r}
# Wiles et al. 2014 data from the Gulf of Alaska ("GOA")
goa.df = read_csv(file.path('data','climateData','wiles_goa.csv'))
goa.mean <- mean(goa.df$temp)
goa.df$yearBP <- 1950- goa.df$yearCE 

goa.df$yearBins <- cut(goa.df$yearBP,include.lowest = T,right = F,
                       breaks = seq(min(goa.df$yearBP),max(goa.df$yearCE),10), 
                       labels = seq(min(goa.df$yearBP),2000,10))
goa.df$zscore = (goa.df$temp - goa.mean) 

goa.binned <- goa.df %>%
  group_by(yearBins) %>%
  summarise(bin.temp = mean(temp)) %>%
  mutate(zscore = (bin.temp - goa.mean)) %>%
  mutate(sign = ifelse(zscore >= 0,'positive','negative'))

# Kaufman et al. pan-Alaska dataset 
clim.df = read_csv(file.path('data','climateData','kaufman_paleoclimate_synthesis.csv'))
clim.df[["tempsign"]] <- ifelse(clim.df[["all.temp"]] >= 0, "positive", "negative")
clim.df[["moistsign"]] <- ifelse(clim.df[["all.moist"]] >= 0, "positive", "negative")
```


## Re-create Figure 5
```{r}
source(file.path('code','plotting_fig_5.r'))
fig_5
```

## Correlation analysis

Read in climate data from Kaufman et al. XXXX, a composite Holocene paleoclimate record for Beringia
```{r}
kaufman_df <- read_csv(file.path('data','climateData','kaufman_paleoclimate_synthesis.csv')) 
```


### Last 10,000 years
Join all of the charcoal-based metrics together with the climate record, binned into 500-year intervals that correspod with Kaufman record.
```{r}
corr_df_10kya <- compositeCHAR %>% 
  filter(region == 'Alaska') %>% 
  right_join(FRIsmooth, by = c('yr.bp' = 'year_BP')) %>%
  right_join(ratio, by = 'yr.bp') %>%
  full_join(pctBurned_ak, by = c('yr.bp' = 'year')) %>%
  mutate(bins500 = as.numeric(as.character(cut(.$yr.bp,include.lowest = T, right = T,
                                            breaks = seq(0,10000,500), 
                                            labels = seq(250,10000,500)))),
         bins500 = ifelse(yr.bp < 0, 250, bins500)) %>%
  group_by(bins500) %>%
  summarise_all(.funs = function(x){mean(x, na.rm = T)} ) %>%
  select(-yr.bp) %>% 
  rename(yr.bp = bins500) %>%
  full_join(kaufman_df, by = 'yr.bp') %>%
  filter(yr.bp <= 8000) %>%
  select(CHAR_500 = composite.mean.500, 
         CHAR_100 = composite.mean.100, 
         Pct_Burned = boot.median, 
         mFRI_1000, 
         ratio,
         all_temp = all.temp,
         all_moist = all.moist,
         midgeA = midgeA.temp,
         midgeB = midgeB.temp,
         pollenAnnual = pollenAnn.temp,
         pollenSummer = pollenSumm.temp) 
```

Calculate Pearson pairwise correlations.
```{r}
Hmisc::rcorr(as.matrix(corr_df_10kya), type="pearson") # type can be pearson or spearman
```

### Last 1200 years
```{r}
wiles_df <- read_csv(file.path('data','climateData','wiles_goa.csv')) %>% 
  mutate(yr.bp = 1950-yearCE)

corr_df_1kya <- wiles_df %>%
  mutate(bins10 = rev(cut(.$yr.bp,include.lowest = T, right = T,
                          breaks = seq(1150,-60,-10),
                          labels = seq(1150,-50,-10)))) %>%
  mutate(bins10 = as.numeric(as.character(bins10))) %>% 
  group_by(bins10) %>%
  summarise_all(.funs = mean) %>%
  select(-yr.bp, -yearCE) %>%
  rename(yr.bp = bins10) %>%
  full_join(.,filter(compositeCHAR, region == 'Alaska')) %>%
  full_join(pctBurned_ak, by = c('yr.bp' = 'year')) %>%
  full_join(FRIsmooth, by = c('yr.bp' = 'year_BP')) %>%
  full_join(ratio) %>% 
  select(CHAR_500 = composite.mean.500, 
         CHAR_100 = composite.mean.100, 
         Pct_Burned = boot.median, 
         mFRI_1000, 
         ratio,
         GOAtemp = temp)
```

Calculate Pearson pairwise correlations.
```{r}
Hmisc::rcorr(as.matrix(corr_df_1kya), type="pearson") # type can be pearson or spearman
```





